---
title: "Almeida-Warren et al. 2022 - Tool site use"
author: "KAW"
date: "24 June 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q2. Tool Site Use Frequency

Descriptive statistics on frequency of tool site use relative to availability of nuts, raw materials, tools, wild food trees, distance to nearest nest cluster and distance to nearest river.

### Project prep

```{r, message=FALSE, warning=FALSE}

# Load libraries

# manipulating dataframes
library(plyr)
library(dplyr)
library(tidyr)

#tables
library(sjPlot) # tab_model
library(flextable) # prep for docx output
library(broom) # glm and drop1 tidy outputs

#plotting
library(ggplot2)
library(ggpubr) # ggarrange combine multiple plots to une figure
library(cowplot) # facet plots with legend
library(Cairo) # image export

```

```{r}
#load data

# Quadrat data
Slc_DF <- read.csv("TS_Selection_DF.csv")

# Tool site data
TS_DF <- read.csv("Tool_Sites.csv")

# Working tools (i.e. stones that are still suitable for nut-cracking)
Tools_DF <- read.csv("NC_AH_tools.csv")

# Data from tool site activity monitoring
TS_act <- read.csv("TS_activity.csv")

```


## Data prep

```{r}
# select activity variables

Act_DF <- TS_act %>% 
  select (c(Week, Q_ID, Recent_activity, Nut_count, Num_regime))

# tally tools by Q_ID (Quadrat ID)
TS_Tools <- Tools_DF %>% group_by (Q_ID) %>% tally() %>% mutate (Tools = n) %>% select (c(Q_ID, Tools))
  
# Join DFs

Act_DF <- Act_DF %>% left_join (TS_DF %>% select(c(Q_ID, Status)), by = "Q_ID")

# subset data to active tool sites (Status = "Active")

Frq_DF <- Act_DF %>% filter (Status == "ACTIVE")

# summarise events by Q_ID

n_events <- Frq_DF %>% filter(Recent_activity == "Y") %>% group_by(Q_ID) %>% 
                     summarise(n_events = n())
n_obs <- Frq_DF %>% drop_na(Recent_activity) %>% group_by(Q_ID) %>% 
                     summarise(n_obs = n())

# summarise Nut_count, Num_regime

Frq_DF <- Frq_DF %>% group_by(Q_ID) %>% summarise(Nut_mean = mean(Nut_count, na.rm = TRUE),
                                                  Reg_mean = mean (Num_regime, na.rm = TRUE))

# Join DFs

Frq_DF <- Frq_DF %>% left_join(n_obs, by = "Q_ID") %>% left_join(n_events, by = "Q_ID")

Frq_DF <- Frq_DF %>% left_join (TS_Tools, by = "Q_ID")

Frq_DF <- Frq_DF %>% left_join (Slc_DF %>% select(c(Q_ID, RM, RM_S3_S5, Wfood_2DBH, Wfood_Fr, Wfood_THV, Nest_dist, River_dist)), by = "Q_ID")


# Remove tool sites that were monitored for less than 10 weeks

Frq_DF <- Frq_DF %>% filter (n_obs > 10)
  
```


```{r}
# Check data distribution

Frq_DF %>% distinct(Q_ID, .keep_all = TRUE) %>% select(n_obs) %>% group_by(n_obs) %>% summarise(n = n())
```

Observations:

- Unequal number of observations per tool site

```{r}
# Scale n_events to n_obs

Frq_DF <- Frq_DF %>% mutate (prop_events = n_events/n_obs)

```


```{r}
# Check data distribution

Frq_DF %>% distinct(Q_ID, .keep_all = TRUE) %>% select(n_events, prop_events) %>% group_by(n_events, prop_events) %>% summarise(n = n())
```

Observations:

- Should recode to 2 categories: <0.15 (n = 10; n_events <= 2) vs. =>0.15 (n = 7; n_events > 2)

```{r}
# Create new variables for the two categorical options

Frq_DF <- Frq_DF %>% mutate(FreqLH = case_when (prop_events < 0.15 ~ "Low",
                                                prop_events >= 0.15 ~ "High"))

```


```{r}
# Reorder variables

Frq_DF <- Frq_DF %>% relocate(c(n_obs, n_events, prop_events, FreqLH), .before = Nut_mean)

```


## Data exploration - Full Data

## Inspect DF

```{r}
str(Frq_DF)
```

DF is composed of 17 observations. Each row represents an individual tool site.

```{r}
#check for missing values

colSums(is.na(Frq_DF))
```

There are no NA's



```{r}
# Summarise data by low vs. high frequency sites

#Low frequency sites

DS_Low <- Frq_DF %>% select(-c(Q_ID:prop_events)) %>% filter(FreqLH == "Low") %>% gather("Variables", "value", 2:10) %>% group_by(Variables) %>% 
  summarise(minL = min(value),
            maxL = max(value),
            meanL = mean(value),
            sdL = sd(value),
            medianL = median(value),
            IQRL = IQR(value))

#High frequency sites

DS_High <- Frq_DF %>% select(-c(Q_ID:prop_events)) %>% filter(FreqLH == "High") %>% gather("Variables", "value", 2:10) %>% group_by(Variables) %>% 
  summarise(minH = min(value),
            maxH = max(value),
            meanH = mean(value),
            sdH = sd(value),
            medianH = median(value),
            IQRH = IQR(value))

# Join DFs

DS_Freq <- DS_Low %>% left_join(DS_High, by = "Variables") %>% mutate_if(is.numeric, round, 3)

DS_Freq

# Create table

TS_Use_DS_Tab <- flextable(DS_Freq)

print(TS_Use_DS_Tab, preview = "docx")

```


## Plots

### overview boxplots

```{r}

#create overview boxplots

Frq_DF %>% group_by (FreqLH) %>% gather("Variables", "value", 6:15) %>% ggplot(aes(as.factor(FreqLH), value)) + 
    geom_boxplot() + geom_jitter(alpha = 0.1) +
    facet_wrap(~Variables, scales = 'free_y')
```


### formatted boxplots using ggarrange

```{r}

#set plot theme

theme_bp <- function(){

theme_classic() + 
    theme(legend.position = "none", panel.grid.major.x = element_blank(),
          axis.text=element_text(size=12), axis.title = element_text(size=14), 
          axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0),
                                      colour = "grey 30"), 
          axis.title.x = element_text(margin = margin(t = 3, r = 0, b = 0, l = 0),
                                      colour = "grey 30"))
  
}
```

```{r}
# Nut mean

BP_Nut <- Frq_DF %>% 
  select(c("FreqLH", "Nut_mean")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Nut_mean, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("\nMean nut availability") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))
  
```

```{r}
# Regime

BP_Regime <- Frq_DF %>% 
  select(c("FreqLH", "Reg_mean")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Reg_mean, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("Mean number\n of regime") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```

```{r}

# RM

BP_RM <- Frq_DF %>% 
  select(c("FreqLH", "RM")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, RM, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("\nRaw materials") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```


```{r}

# RM_S3_S5

BP_RM35 <- Frq_DF %>% 
  select(c("FreqLH", "RM_S3_S5")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, RM_S3_S5, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("Raw materials\n (size class 3 - 5)") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```

```{r}

# Tools

BP_Tools <- Frq_DF %>% 
  select(c("FreqLH", "Tools")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Tools, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("\nTools") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```


```{r}

# Food trees

BP_FT <- Frq_DF %>% 
  select(c("FreqLH", "Wfood_2DBH")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Wfood_2DBH, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("\nWild food trees") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))


```

```{r}

#Fr trees

BP_FrT <- Frq_DF %>% 
  select(c("FreqLH", "Wfood_Fr")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Wfood_Fr, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  xlab("Frequency of events") + ylab("\nWild fruit trees") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```

```{r}

# Nest_dist

BP_Nest <- Frq_DF %>% 
  select(c("FreqLH", "Nest_dist")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, Nest_dist, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  scale_y_continuous(breaks = c(100,200,300,400,500), labels = c(0.1, 0.2, 0.3, 0.4, 0.5)) +
  xlab("Frequency of events") + ylab("Distance to nearest\n nest cluster (km)") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```

```{r}

# River_dist

BP_River <- Frq_DF %>% 
  select(c("FreqLH", "River_dist")) %>%
  group_by (FreqLH) %>%
  ggplot(aes(FreqLH, River_dist, fill=FreqLH)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.2, size = 4) +
  stat_summary(fun=mean, colour="black", geom="point", shape=18, size=4) +
  theme_bp() +
  scale_x_discrete(limits=c("Low", "High")) +
  scale_y_continuous(breaks = c(100,200,300), labels = c(0.1, 0.2, 0.3)) +
  xlab("Frequency of events") + ylab("Distance to\n nearest river (km)") +
  scale_fill_grey(start = 1, end = 0.4)  +
  theme(plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))

```

```{r}

# Combine boxplots of 6 main variables

BP_R1 <- ggarrange (BP_Tools, BP_RM35, BP_Nut, 
           ncol = 3, labels = c("a", "b", "c"),
           legend = NULL,
           align = "hv",
           hjust = -2.8, vjust = 2)

BP_R2 <- ggarrange (BP_FT, BP_Nest, BP_River,
           ncol = 3, labels = c("d", "e", "f"),
                   legend = NULL,
                       align = "hv",
                       hjust = -2.8, vjust = 2)

BP_ReUse <- ggarrange (BP_R1, BP_R2,
           nrow = 2, align = "hv")

# Save

ggsave("TS_FreqUse_BP.tiff", BP_ReUse, type = "cairo", scale = 3, width = 3, height = 2, dpi = 300)

BP_ReUse

```
