---
title: "Almeida-Warren et al. 2022 - Tool site inactivity"
author: "KAW"
date: "24 June 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q3. Tool Site Inactivity

Analysis of tool site inactivity at the Quadrat level using binomial GLM (response: abandoned(1)/active(0) nut-cracking site; predictors of interest: prod nut tree, raw materials, tools, wild food trees)

### Project prep

```{r, message=FALSE, warning=FALSE}
# Load libraries

# manipulating dataframes
library(plyr)
library(dplyr)
library(tidyr)

# model exploration
library(lattice) # cleveland dotplot
library(corrplot) # correlation
library(car) # vif


# model stats
library(MuMIn) # pseudo-R2; AICc

#tables
library(sjPlot) # tab_model
library(flextable) # prep for docx output
library(broom) # glm and drop1 tidy outputs

#plotting
library(ggplot2)
library(ggeffects) # ggpredict
library(visreg) # glm models
library(ggpubr) # ggarrange combine multiple plots to une figure
library(cowplot) # facet plots with legend
library(Cairo) # image export

```

```{r }
#load data

# Quadrat data
Slc_DF <- read.csv("TS_Selection_DF.csv")

# Tool site data
TS_DF <- read.csv("Tool_Sites.csv")

# Working tools (i.e. stones that are still suitable for nut-cracking)
Tools_DF <- read.csv("NC_AH_tools.csv")

# Data from tool site monitoring
Monitoring_DF <- read.csv("TS_monitoring_DF.csv")


```

```{r}
#Define plot theme

theme_BMS <- function(){
  
  theme_light() + theme(legend.text = element_text (size = 14), 
                     legend.title = element_text(size = 16), 
                     panel.grid = element_blank(), 
                     axis.text = element_text(size=12, colour = "grey 20"), 
                     axis.title = element_text(size=14), 
                     axis.title.y = element_text(margin = margin(t = 0, r = 10, 
                                                                 b = 0, l = 0),
                                                 colour = "grey 20"), 
                     axis.title.x = element_text(margin = margin(t = 3, r = 0, 
                                                                 b = 0, l = 0),
                                                 colour = "grey 20"),
                     axis.ticks = element_line(colour = "grey 20", size = 0.5),
                     panel.border = element_rect(colour = "grey 20", fill=NA, size = 1),
                     strip.text.x = element_text(size = 18, colour = NULL, 
                                                 angle = 0),
                     strip.text.y = element_text(size = 18, colour = NULL, 
                                                 angle = 0))
}



```

## Data prep

```{r}
# select Quadrat variables

Abn_DF <- Slc_DF %>% 
  filter (TS_Pres == 1) %>%
  select (c(Q_ID, RM, RM_S3_S5, Wfood_2DBH, Wfood_Fr))
  
# tally tools by Q_ID
TS_Tools <- Tools_DF %>% group_by (Q_ID) %>% tally() %>% mutate (Tools = n) %>% select (c(Q_ID, Tools))

# summarise Nut availability data

Nut_availability <- Monitoring_DF %>%
                     drop_na(Nut_count) %>%
                     group_by(Q_ID) %>%         
                     summarise(N_obs = n(),
                               Nut_mean = mean(Nut_count),
                               Nut_sd = sd(Nut_count),
                               Nut_median = median (Nut_count),
                               Nut_max = max (Nut_count))

# Join DFs
Abn_DF <- Abn_DF %>% left_join (TS_Tools, by = "Q_ID")

Abn_DF <- Abn_DF %>% left_join (TS_DF %>% select(c(Q_ID, Status)), by = "Q_ID")

Abn_DF <- Abn_DF %>% left_join (Nut_availability %>% select(c(Q_ID, N_obs, Nut_mean, Nut_median, Nut_max)), by = "Q_ID")

# Recode Status to binomial variable

Abn_DF <- Abn_DF %>% mutate(Inactive = case_when (Status == "ACTIVE" ~ 0,
                                                   Status == "INACTIVE" ~ 1))
  
```


```{r}
# Check data distribution

Abn_DF %>% select(Inactive) %>% group_by(Inactive) %>% summarise(n = n())
```

Observations:

- 24 (~60%) active sites; 16 (~40%) Inactive sites

```{r}
# Reorder variables

Abn_DF <- Abn_DF %>% relocate(c(Inactive, Nut_mean, Nut_median, Nut_max), .before = RM)
```



## Data exploration - Full Data

## Inspect DF

```{r}
str(Abn_DF)
```

DF is composed of 40 observations. Each row represents an individual Quadrat. Inactive is binomial and is coded as 0 (active) and 1 (inactive); All other variables are quantitative.

```{r}
#check for missing values

colSums(is.na(Abn_DF))
```

There are no NA's

## Check for outliers

### Barplots

```{r}

Abn_DF %>% select (-Status) %>% group_by (Inactive) %>% gather("Variables", "value", 3:10) %>% ggplot(aes(as.factor(Inactive), value)) + 
    geom_boxplot() + geom_jitter(alpha = 0.1) +
    facet_wrap(~Variables, scales = 'free_y')

```

Observations:

- Distributions look reasonable; possible outlier in RM, RM_S3_S5

### Cleveland dotplots

```{r}
# lattice version

dotplot(as.matrix(Abn_DF[,c(3:10)]), groups = FALSE,
        strip = strip.custom(bg = 'white',
        par.strip.text = list(cex = 0.8)),
       scales = list(x = list(relation = "free",
                              draw = TRUE),
                     y = list(relation = "free", 
                              draw = FALSE)),
        col = 1, cex  = 0.5, pch = 16,
        xlab = "Data range",
        ylab = "Order of observations")

```


Observations:

- Visible outlier in RM; RM_S3_S5


Check of notes and original data suggests outliers are real observations:

- RM/RM_S3_S5 - high value @ Obs 11 (TS19_Q0) & 28 (TS34_Q0) due to human-made stone pile resulting from historical forest clearing for planting crops;


## Check data distribution

```{r}

Abn_DF %>% select (c(-1)) %>% gather("Variables", "value", 1:9) %>% ggplot(aes(value)) + 
    geom_histogram() + 
    facet_wrap(~Variables, scales = "free")

```


Observations:

- Single outliers in RM, RM_S3_S5 & Tools


## Check multicollinearity among covatiates

### Pairplots

```{r}

# Creat DF of explanatory variables

Abn_ExpVar <- Abn_DF %>% dplyr::select (-c(Q_ID, Inactive, Status, N_obs))

# Correlation panels

panel.cor.pearson <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y, method = "pearson"), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 1/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * abs(r))
}

panel.cor.spearman <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y, method = "spearman"), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 1/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * abs(r))
}

# Pairplots

## pearson correlation

PairPlots_AbnP <- pairs (Abn_ExpVar,
                        lower.panel = panel.cor.pearson)

PairPlots_AbnP


## spearman correlation

PairPlots_AbnS <- pairs (Abn_ExpVar,
                        lower.panel = panel.cor.spearman)

PairPlots_AbnS

```


"(...)Spearman rank correlation makes no assumptions about linearity in the relationship between the two variables (Zar, 1996).Booth et al. (1994) suggest that correlations between pairs of variables with magnitudes greater than ±0.5 indicate high collinearity, and we use this rough rule-of-thumb here." (Zuur et al. 2009, p. 473)

Observations:

Excluding related pairs of variables (Nut_ ~ Nut_; RM ~ RM_; RM(_) ~ Tools; Wfood_2DBH ~ Wfood_Fr):

- Pearson cor range: [-0.13, 0.21] (Under ±0.5)
- Spearman cor range: [-0.13, 0.28] (Under ±0.5)


CONCLUSIONS FROM DATA EXPLORATION:

- No NA's
- Potential outlier in Tools and RM_ variables
- No collinearity issues between unrelated variables
- Overall, data exploration did not reveal any unusual patters


#Variable Selection

## Nut_ variables

Nut_mean; Nut_median; Nut_max -> Choosing Nut_mean, because data distribution is more normal

## RM, RM_S3_S5, Tools

- RM as default variable
- Fit 1 other models with RM replaced by RM_S3_S5
- Fit 1 other model with RM replaced by Tools

## Wfood_2DBH; Wfood_Fr

- Choosing Wfood_DBH as default variable
- Fit a model with Wfood_2DBH replaced by Wfood_Fr

#Model 1

binomial GLM (response: yes/no inactive; predictors: mean nut availability, raw materials, wild food trees)


### Variance inflation factors

```{r}
# Variance inflation factors

vif(glm(Inactive ~ Nut_mean + RM + Wfood_2DBH, data = Abn_DF))

```

Observations:

- VIF ~1 for all variables; low co-variance.
- All covariates can be used in the model


```{r}
Abn_GLM_M1 <- glm(as.factor(Inactive) ~ Nut_mean + RM + Wfood_2DBH, data = Abn_DF, family = binomial(link = "logit"))

summary(Abn_GLM_M1)

```

## Compare with null model

```{r}
# Null model

Abn_GLM_M1_Null <- glm(as.factor(Inactive) ~ 1, data = Abn_DF, family = binomial(link = "logit"))


# anova Null vs. Full model

AD_M1 <- anova(Abn_GLM_M1_Null, Abn_GLM_M1, test = "LRT")

AD_M1
```

Observations:

p < 0.05; ONLY JUST! - can reject null model.


## Likelihood ratio test of main effects

```{r}
# drop1

d1_M1 <- drop1(Abn_GLM_M1, test = "Chisq")

d1_M1
```

### Make table with drop1 results

```{r}
# create df from glm results

Tab_M1 <- tidy(Abn_GLM_M1)

# replace statistic and p with values from drop1

Tab_M1$statistic <- d1_M1$LRT

Tab_M1$p.value <- d1_M1$`Pr(>Chi)`

Tab_M1 <- Tab_M1 %>% mutate_if(is.numeric, round, 3)

Tab_M1

#create table

Abn_Tab_M1 <- flextable(Tab_M1)

print(Abn_Tab_M1, preview = "docx")

```

# Model 1a

binomial GLM (response: yes/no inactive; predictors: mean nut availability, raw materials (size class 3-5), wild food trees)

## Variance inflation factors

```{r}
# Variance inflation factors

vif(glm(Inactive ~ Nut_mean + RM_S3_S5 + Wfood_2DBH, data = Abn_DF))
```


Observations:
-  < 1.1 for all variables; low co-variance.
- All covariates can be used in the model


## Fit M1a

```{r}
Abn_GLM_M1a <- glm(as.factor(Inactive) ~ Nut_mean + RM_S3_S5 + Wfood_2DBH, data = Abn_DF, family = binomial(link = "logit"))

summary(Abn_GLM_M1a)
```

## Compare with null model

```{r}
# Null Model

Abn_GLM_M1a_Null <- glm(as.factor(Inactive) ~ 1, family = binomial(link = "logit"), data = Abn_DF)

# anova Null vs. Full model

AD_M1a <- anova(Abn_GLM_M1a_Null, Abn_GLM_M1a, test = "LRT")

AD_M1a
```

Observations:

p > 0.05 - cannot reject null model.

## Likelihood ratio test of main effects

```{r}
# drop1

d1_M1a <- drop1(Abn_GLM_M1a, test = "Chisq")

d1_M1a
```

## Summary Table

### Make table with drop1 results

```{r}
# create df from glm results

Tab_M1a <- tidy(Abn_GLM_M1a)

# replace statistic and p with values from drop1

Tab_M1a$statistic <- d1_M1a$LRT

Tab_M1a$p.value <- d1_M1a$`Pr(>Chi)`

Tab_M1a <- Tab_M1a %>% mutate_if(is.numeric, round, 3)

Tab_M1a

#create table

Abn_Tab_M1a <- flextable(Tab_M1a)

print(Abn_Tab_M1a, preview = "docx")

```



#Model 1a

binomial GLM (response: yes/no inactive; predictors: mean nut availability, tools, wild food trees)


### Variance inflation factors

```{r}
# Variance inflation factors

vif(glm(Inactive ~ Nut_mean + Tools + Wfood_2DBH, data = Abn_DF))

```

Observations:

- VIF ~1 for all variables; low co-variance.
- All covariates can be used in the model


```{r}
Abn_GLM_M1b <- glm(as.factor(Inactive) ~ Nut_mean + Tools + Wfood_2DBH, data = Abn_DF, family = binomial(link = "logit"))

summary(Abn_GLM_M1b)

```

## Compare with null model

```{r}

# anova Null vs. Full model

AD_M1b <- anova(Abn_GLM_M1_Null, Abn_GLM_M1b, test = "LRT")

AD_M1b
```

Observations:

p < 0.05; can reject null model.


## Likelihood ratio test of main effects

```{r}
# drop1

d1_M1b <- drop1(Abn_GLM_M1b, test = "Chisq")

d1_M1b
```

## Graphical validation

### Homogeneity of variance 

```{r}
#Pearson vs. fitted residuals

Fitted_M1b <- fitted(Abn_GLM_M1b)
Resid_M1b <- resid(Abn_GLM_M1b, type = "pearson")
par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.2)
plot(x = Fitted_M1b, y = Resid_M1b,
xlab = "Fitted values",
ylab = "Pearson Residuals")
abline(h = 0, lty = 2)
```

Observations:

- Distribution of residuals is consistent along the horizontal axis

### Model misfit

```{r}
# Pearson residuals vs. covariates

#Nut_mean

plot(x = Abn_DF$Nut_mean,
y = Resid_M1b,
xlab = "Nut_mean",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```

```{r}
#Tools

plot(x = Abn_DF$Tools,
y = Resid_M1b,
xlab = "Tools",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```


```{r}
#WFood_2DBH

plot(x = Abn_DF$Wfood_2DBH,
y = Resid_M1b,
xlab = "Wfood_2DBH",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```

Observations:

- Distributions for all covariates show no obvious patterns for the residuals, except for one extreme value

### Absence of influentia observations

```{r}
#Cook's distance

par(mfrow = c(1, 1))
plot(cooks.distance(Abn_GLM_M1b),
xlab = "Observation",
ylab = "Cook's distance",
type = "h",
ylim = c(0, 1.1),
cex.lab = 1.5)
abline(h = 1, lty = 2)
```

"Cook distance larger than 1 (...) is the threshold value upon [which] one should take further action (Fox, 2002)" (Zuur et al. 2009, p. 27)

Observations:

- Max CD < 0.5; No evidence of influential observations
- Model validation has shown no evidence of model misfit, model residuals are acceptable and there are no influential observations


## Summary Table

### Make table with drop1 results

```{r}
# create df from glm results

Tab_M1b <- tidy(Abn_GLM_M1b)

# replace statistic and p with values from drop1

Tab_M1b$statistic <- d1_M1b$LRT

Tab_M1b$p.value <- d1_M1b$`Pr(>Chi)`

Tab_M1b <- Tab_M1b %>% mutate_if(is.numeric, round, 3)

#create table

Abn_Tab_M1b <- flextable(Tab_M1b)

print(Abn_Tab_M1b, preview = "docx")

```

#Model 1b

binomial GLM (response: yes/no inactive; predictors: mean nut availability, raw materials (size class 3 - 5), wild fruit trees)


### Variance inflation factors

```{r}
# Variance inflation factors

vif(glm(Inactive ~ Nut_mean + RM_S3_S5 + Wfood_Fr, data = Abn_DF))

```

Observations:

- VIF <1.1 for all variables; low co-variance.
- All covariates can be used in the model


```{r}
Abn_GLM_M1c <- glm(as.factor(Inactive) ~ Nut_mean + RM_S3_S5 + Wfood_Fr, data = Abn_DF, family = binomial(link = "logit"))

summary(Abn_GLM_M1c)

```

```{r}

# anova Null vs. Full model

AD_M1c <- anova(Abn_GLM_M1_Null, Abn_GLM_M1c, test = "LRT")

AD_M1c
```

Observations:

p > 0.05 - cannot reject null model


## Likelihood ratio test of main effects

```{r}
# drop1

d1_M1c <- drop1(Abn_GLM_M1c, test = "Chisq")

d1_M1c
```

## Summary Table

### Make table with drop1 results

```{r}
# create df from glm results

Tab_M1c <- tidy(Abn_GLM_M1c)

# replace statistic and p with values from drop1

Tab_M1c$statistic <- d1_M1c$LRT

Tab_M1c$p.value <- d1_M1c$`Pr(>Chi)`

Tab_M1c <- Tab_M1c %>% mutate_if(is.numeric, round, 3)

Tab_M1c

#create table

Abn_Tab_M1c <- flextable(Tab_M1c)

print(Abn_Tab_M1c, preview = "docx")

```

#Model 1c

binomial GLM (response: yes/no inactive; predictors: mean nut availability, tools, wild fruit trees)


### Variance inflation factors

```{r}
# Variance inflation factors

vif(glm(Inactive ~ Nut_mean + Tools + Wfood_Fr, data = Abn_DF))

```

Observations:

- VIF <1.1 for all variables; low co-variance.
- All covariates can be used in the model


```{r}
Abn_GLM_M1d <- glm(as.factor(Inactive) ~ Nut_mean + Tools + Wfood_Fr, data = Abn_DF, family = binomial(link = "logit"))

summary(Abn_GLM_M1d)

```

```{r}

# anova Null vs. Full model

AD_M1d <- anova(Abn_GLM_M1_Null, Abn_GLM_M1d, test = "LRT")

AD_M1d
```

Observations:

p > 0.05. Can reject Null model

## Likelihood ratio test of main effects

```{r}
# drop1

d1_M1d <- drop1(Abn_GLM_M1d, test = "Chisq")

d1_M1d
```

## Graphical validation

### Homogeneity of variance 

```{r}
#Pearson vs. fitted residuals

Fitted_M1d <- fitted(Abn_GLM_M1d)
Resid_M1d <- resid(Abn_GLM_M1d, type = "pearson")
par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.2)
plot(x = Fitted_M1d, y = Resid_M1d,
xlab = "Fitted values",
ylab = "Pearson Residuals")
abline(h = 0, lty = 2)
```

Observations:

- Distribution of residuals is consistent along the horizontal axis

### Model misfit

```{r}
# Pearson residuals vs. covariates

#Nut_mean

plot(x = Abn_DF$Nut_mean,
y = Resid_M1d,
xlab = "Nut_mean",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```

```{r}
#Tools

plot(x = Abn_DF$Tools,
y = Resid_M1d,
xlab = "Tools",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```


```{r}
#WFood_2DBH

plot(x = Abn_DF$Wfood_2DBH,
y = Resid_M1d,
xlab = "Wfood_2DBH",
ylab = "Pearson residuals",
pch = 16, cex = 1.5)
abline(h = 0, lty = 2)

```

Observations:

- Distributions for all covariates show no obvious patterns for the residuals, except for one extreme value

### Absence of influentia observations

```{r}
#Cook's distance

par(mfrow = c(1, 1))
plot(cooks.distance(Abn_GLM_M1d),
xlab = "Observation",
ylab = "Cook's distance",
type = "h",
ylim = c(0, 1.1),
cex.lab = 1.5)
abline(h = 1, lty = 2)
```

"Cook distance larger than 1 (...) is the threshold value upon [which] one should take further action (Fox, 2002)" (Zuur et al. 2009, p. 27)

Observations:

- Max CD < 0.7; No evidence of influential observations
- Model validation has shown no evidence of model misfit, model residuals are acceptable and there are no influential observations

## Summary Table

### Make table with drop1 results

```{r}
# create df from glm results

Tab_M1d <- tidy(Abn_GLM_M1d)

# replace statistic and p with values from drop1

Tab_M1d$statistic <- d1_M1d$LRT

Tab_M1d$p.value <- d1_M1d$`Pr(>Chi)`

Tab_M1d <- Tab_M1d %>% mutate_if(is.numeric, round, 3)

Tab_M1d

#create table

Abn_Tab_M1d <- flextable(Tab_M1d)

print(Abn_Tab_M1d, preview = "docx")

```

#Model 1d


# Compare models

## Full-Null summary

```{r}
# ANOVA

Abn_ANOVA <- bind_rows("Abn_GLM_M1" = tidy(AD_M1), "Abn_GLM_M1a" = tidy(AD_M1a), "Abn_GLM_M1b" = tidy(AD_M1b), "Abn_GLM_M1c" = tidy(AD_M1c), "Abn_GLM_M1d" = tidy(AD_M1d), .id = "Model")

Abn_ANOVA <- Abn_ANOVA %>% mutate_if(is.numeric, round, 3)

Abn_Tab_ANOVA <- flextable(Abn_ANOVA)

print(Abn_Tab_ANOVA, preview = "docx")

Abn_ANOVA 

```

## AIC for small sample size 

[see Burnham and Anderson, 2002]

```{r}
# AICc

Abn_AIC <- AICc(Abn_GLM_M1, Abn_GLM_M1a, Abn_GLM_M1b, Abn_GLM_M1c, Abn_GLM_M1d)

Abn_AIC <- Abn_AIC %>% mutate(Model = rownames(Abn_AIC)) %>% relocate(Model, .before = "df") %>% mutate_if(is.numeric, round, 3)

Abn_Tab_AIC <- flextable(Abn_AIC)

print(Abn_Tab_AIC, preview = "docx")

```

Abn_GLM_M1a has the lowest AICc - selected as the final model

# Final Model - M1c 


## Plots

### M1c

```{r}
# Nut_mean

Pl_M1b_Nut<- visreg(Abn_GLM_M1b, "Nut_mean", scale = "response",
                      partial=FALSE, jitter = TRUE, alpha=0.1, 
                      line = list (col = "grey20"), 
                      fill = list (fill = "grey90"), 
                      gg = TRUE) +
                geom_point(colour = "grey20", shape = 16) + theme_BMS() + 
                xlab("Mean nut availability") + ylab("Odds of tool site inactivity") +
                scale_x_continuous()


Pl_M1b_Nut

```

```{r}
# Tools

Pl_M1b_Tools <- visreg(Abn_GLM_M1b, "Tools", scale = "response",
                        partial=FALSE, jitter = TRUE, alpha=0.1, 
                        line = list (col = "grey20"), 
                        fill = list (fill = "grey90"),
                        gg = TRUE) +
                  geom_point(colour = "grey20", shape = 16) + theme_BMS() + 
                  xlab("Number of tools") + ylab("Odds of tool site inactivity") +
                  scale_x_continuous(breaks = c(0,5,10,15,20,25))

Pl_M1b_Tools

```


```{r}
# Facet plots

Pl_M1b <- ggarrange(Pl_M1b_Nut, Pl_M1b_Tools,
                   ncol = 2, labels = c("a", "b"),
                   legend = NULL,
                       align = "h",
                       hjust = -0.8)

Pl_M1b

ggsave("TS_Abn_M1b.tiff", Pl_M1b, type = "cairo", scale = 3, width = 2.5, height = 1.5, dpi = 300)
```



